---
title: "Cleaning Census Data"
author: "Team Voteen"
date: "10/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Initial set-up

```{r, message = FALSE}
## Clear environment
rm(list=ls(all=TRUE))

## Set working directory
setwd("~/Documents/GitHub/prop18/data-challenge/01-raw-data/census")

## Load packages
library(tidyverse)

## Import data (skip command skips the first n rows in the file)
census_2018 <- read.csv("voting-age-state-2018.csv", 
                        stringsAsFactors = FALSE, skip = 4)

census_2016 <- read.csv("voting-age-state-2016.csv", 
                        stringsAsFactors = FALSE, skip = 4)

census_2014 <- read.csv("voting-age-state-2014.csv", 
                        stringsAsFactors = FALSE, skip = 3)

census_2012 <- read.csv("voting-age-state-2012.csv", 
                        stringsAsFactors = FALSE, skip = 3)

census_2010 <- read.csv("voting-age-state-2010.csv", 
                        stringsAsFactors = FALSE, skip = 2)
```

## Clean each election year data

The data from the U.S. Census Bureau on age groups by state was available for 
download as individual Excel files. We cleaned the data for each year and then 
combine them using `rbind()`.

We show how we cleaned one year of data and repeat the process for all years.

## 2018:

```{r, message=FALSE}
## Select variables of interest by column number 
census_2018 <- census_2018[c(1,2,4,5,6,8,10,11,13)]

## Rename variables by indexing column number (naming appears strange due to
# excel file formatting)
names(census_2018)[1] <- "state"
names(census_2018)[2] <- "age"
names(census_2018)[3] <- "total_cit"
names(census_2018)[4] <- "total_register"
names(census_2018)[5] <- "percent_register"
names(census_2018)[6] <- "percent_register_cit"
names(census_2018)[7] <- "total_vote"
names(census_2018)[8] <- "percent_vote"
names(census_2018)[9] <- "percent_vote_cit"

## Create a variable for year
census_2018 <- census_2018 %>%
  mutate(year=2018)

## Remove commas from characters to convert into numeric values
census_2018[] <- lapply(census_2018, gsub, pattern = ',', replacement = '')

## Change to numeric values
cols.num <- c("total_cit", "total_register", "percent_register", "percent_register_cit",
              "total_vote", "percent_vote", "percent_vote_cit")

census_2018[cols.num] <- sapply(census_2018[cols.num], as.numeric)

sapply(census_2018, class) # check

## Remove unnecessary rows
census_2018 <- census_2018[-which(census_2018$age == ""), ]

## Change age groups into factors
census_2018$age <- as.factor(census_2018$age)

class(census_2018$age)

## Loop through data to create state names: go through each row of data and 
# insert the value of the row above it (to fill in blanks with state names)
for(i in 1:nrow(census_2018)){
  if(census_2018$state[i] == "") {
    census_2018$state[i] = census_2018$state[i-1]
  }
}

# Show the data is cleaned
head(census_2018, 5)
```

## 2016:

```{r}
census_2016 <- census_2016[c(1,2,4,5,6,8,10,11,13)]

names(census_2016)[1] <- "state"
names(census_2016)[2] <- "age"
names(census_2016)[3] <- "total_cit"
names(census_2016)[4] <- "total_register"
names(census_2016)[5] <- "percent_register"
names(census_2016)[6] <- "percent_register_cit"
names(census_2016)[7] <- "total_vote"
names(census_2016)[8] <- "percent_vote"
names(census_2016)[9] <- "percent_vote_cit"

census_2016 <- census_2016 %>%
  mutate(year=2016)

census_2016[] <- lapply(census_2016, gsub, pattern = ',', replacement = '')

cols.num <- c("total_cit", "total_register", "percent_register", "percent_register_cit",
              "total_vote", "percent_vote", "percent_vote_cit")

census_2016[cols.num] <- sapply(census_2016[cols.num], as.numeric)

sapply(census_2016, class) # check

census_2016 <- census_2016[-which(census_2016$age == ""), ]

census_2016$age <- as.factor(census_2016$age)

class(census_2016$age)

for(i in 1:nrow(census_2016)){
  if(census_2016$state[i] == "") {
    census_2016$state[i] = census_2016$state[i-1]
  }
}

head(census_2016, 5)
```

## 2014: 

```{r}
census_2014 <- census_2014[c(1,2,4,5,6,8,10,11,13)]

names(census_2014)[1] <- "state"
names(census_2014)[2] <- "age"
names(census_2014)[3] <- "total_cit"
names(census_2014)[4] <- "total_register"
names(census_2014)[5] <- "percent_register"
names(census_2014)[6] <- "percent_register_cit"
names(census_2014)[7] <- "total_vote"
names(census_2014)[8] <- "percent_vote"
names(census_2014)[9] <- "percent_vote_cit"

census_2014 <- census_2014 %>%
  mutate(year=2014)

census_2014[] <- lapply(census_2014, gsub, pattern = ',', replacement = '')

cols.num <- c("total_cit", "total_register", "percent_register", "percent_register_cit",
              "total_vote", "percent_vote", "percent_vote_cit")

census_2014[cols.num] <- sapply(census_2014[cols.num], as.numeric)

sapply(census_2014, class) # check

census_2014 <- census_2014[-which(census_2014$age == ""), ]

census_2014$age <- as.factor(census_2014$age)

class(census_2014$age)

for(i in 1:nrow(census_2014)){
  if(census_2014$state[i] == "") {
    census_2014$state[i] = census_2014$state[i-1]
  }
}

head(census_2014, 5)
```

## 2012:

```{r}
census_2012 <- census_2012[c(1,2,4,5,6,8,10,11,13)]

names(census_2012)[1] <- "state"
names(census_2012)[2] <- "age"
names(census_2012)[3] <- "total_cit"
names(census_2012)[4] <- "total_register"
names(census_2012)[5] <- "percent_register"
names(census_2012)[6] <- "percent_register_cit"
names(census_2012)[7] <- "total_vote"
names(census_2012)[8] <- "percent_vote"
names(census_2012)[9] <- "percent_vote_cit"

census_2012 <- census_2012 %>%
  mutate(year=2012)

census_2012[] <- lapply(census_2012, gsub, pattern = ',', replacement = '')

cols.num <- c("total_cit", "total_register", "percent_register", "percent_register_cit",
              "total_vote", "percent_vote", "percent_vote_cit")

census_2012[cols.num] <- sapply(census_2012[cols.num], as.numeric)

sapply(census_2012, class) # check

census_2012 <- census_2012[-which(census_2012$age == ""), ]

census_2012$age <- as.factor(census_2012$age)

class(census_2012$age)

for(i in 1:nrow(census_2012)){
  if(census_2012$state[i] == "") {
    census_2012$state[i] = census_2012$state[i-1]
  }
}

head(census_2012, 5)
```

## 2010:

```{r}
census_2010 <- census_2010[c(1,2,4,5,6,8,10,11,13)]

names(census_2010)[1] <- "state"
names(census_2010)[2] <- "age"
names(census_2010)[3] <- "total_cit"
names(census_2010)[4] <- "total_register"
names(census_2010)[5] <- "percent_register"
names(census_2010)[6] <- "percent_register_cit"
names(census_2010)[7] <- "total_vote"
names(census_2010)[8] <- "percent_vote"
names(census_2010)[9] <- "percent_vote_cit"

census_2010 <- census_2010 %>%
  mutate(year=2010)

census_2010[] <- lapply(census_2010, gsub, pattern = ',', replacement = '')

cols.num <- c("total_cit", "total_register", "percent_register", "percent_register_cit",
              "total_vote", "percent_vote", "percent_vote_cit")

census_2010[cols.num] <- sapply(census_2010[cols.num], as.numeric)

sapply(census_2010, class) # check

census_2010 <- census_2010[-which(census_2010$age == ""), ]

census_2010$age <- as.factor(census_2010$age)

class(census_2010$age)

for(i in 1:nrow(census_2010)){
  if(census_2010$state[i] == "") {
    census_2010$state[i] = census_2010$state[i-1]
  }
}

head(census_2010, 5)
```

## Combine all years into one dataset

```{r}
census_vote <- rbind(census_2018, census_2016, census_2014, census_2012, census_2010)

head(census_vote)

table(census_vote$year)
```

## Write to a csv file and save to folder

**Note:** User can specify the file path location where the file is saved and 
the name of the file ("name.csv").
```{r}
write.csv(census_vote,
          "~/Documents/GitHub/prop18/data-challenge/02-clean-data/clean_census_vote.csv")
```




