---
title: "Create choropleth with youth vote data"
author: "Team Voteen"
date: "10/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Initial set-up
```{r, message=FALSE}
## Clear environment
rm(list=ls(all=TRUE))

## Set working directory
setwd("~/Documents/GitHub/prop18/data-challenge/02-clean-data")

## Load packages
library(tidyverse)
library(leaflet)
library(tigris)

## Import data
youth_vote <- read.csv("clean_youth_vote_200928.csv")
```

## Make shape files and polygons using `tigris` package (includes pre-loaded map of U.S.) and `leaflet`
```{r, message = FALSE, echo = FALSE}
# Import polygon from shape file
states <- states(cb = T)

# Tiles for US states
states %>% 
  leaflet() %>% 
  addTiles() %>% 
  addPolygons(popup = ~NAME)
```

## Subset data by election year
```{r}
## Split dataset by year
youth_vote_16 <- filter(youth_vote, year == 2016)

youth_vote_18 <- filter(youth_vote, year == 2018)
```

## Merge geospatial data (with coordinates) with youth vote data
```{r}
youth_vote_16 <- geo_join(states, youth_vote_16, "STUSPS", "state_code")

youth_vote_18 <- geo_join(states, youth_vote_18, "STUSPS", "state_code")
```

## Convert primary voting variable into factor variable
```{r}
youth_vote_16$primary_17 <- ifelse(youth_vote_16$primary_17 %in% "Y", "Yes", "No")
youth_vote_16$primary_17 <- as.factor(youth_vote_16$primary_17)

youth_vote_18$primary_17 <- ifelse(youth_vote_18$primary_17 %in% "Y", "Yes", "No")
youth_vote_18$primary_17 <- as.factor(youth_vote_18$primary_17)
```

## Define color palettes for different variables
```{r}
pal_turnout <- colorFactor("Purples", domain = youth_vote_16$primary_17)
pal_register <- colorFactor("Oranges", domain = youth_vote_18$primary_17)
```

## Format pop-up text
```{r}
popup_turnout_16 <- paste0(youth_vote_16$state, "<br>Youth Turnout (2016): ", 
                        as.character(youth_vote_16$turnout * 100), "%")
popup_turnout_18 <- paste0(youth_vote_18$state, "<br>Youth Turnout (2018): ", 
                           as.character(youth_vote_18$turnout * 100), "%")
popup_register_16 <- paste0(youth_vote_16$state, "<br>Youth Registration (2016): ", 
                           as.character(youth_vote_16$register * 100), "%")
popup_register_18 <- paste0(youth_vote_18$state, "<br>Youth Registration (2018): ", 
                           as.character(youth_vote_18$register * 100), "%")
```

## 17-year-old Primary voting and youth voter turnout
```{r}
# Generate map for 2016
leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(-98.483330, 38.712046, zoom = 4) %>% 
  addPolygons(data = youth_vote_16, 
              fillColor = ~pal_turnout(youth_vote_16$primary_17), 
              fillOpacity = 0.7, 
              weight = 0.2, 
              smoothFactor = 0.2,
              popup = ~popup_turnout_16) %>%
  addLegend(pal = pal_turnout,
            values = youth_vote_16$primary_17,
            position = "bottomright", 
            title = "States with Primary Voting<br>for 17 year-olds")
```

```{r}
# Generate map for 2018
leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(-98.483330, 38.712046, zoom = 4) %>% 
  addPolygons(data = youth_vote_18, 
              fillColor = ~pal_turnout(youth_vote_18$primary_17), 
              fillOpacity = 0.7, 
              weight = 0.2, 
              smoothFactor = 0.2,
              popup = ~popup_turnout_18) %>%
  addLegend(pal = pal_turnout,
            values = youth_vote_18$primary_17,
            position = "bottomright", 
            title = "States with Primary Voting<br>for 17 year-olds")
```

## 17-year-old Primary voting and youth voter registration
```{r}
# Generate map for 2016
leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(-98.483330, 38.712046, zoom = 4) %>% 
  addPolygons(data = youth_vote_16, 
              fillColor = ~pal_register(youth_vote_16$primary_17), 
              fillOpacity = 0.7, 
              weight = 0.2, 
              smoothFactor = 0.2,
              popup = ~popup_register_16) %>%
  addLegend(pal = pal_register,
            values = youth_vote_16$primary_17,
            position = "bottomright", 
            title = "States with Primary Voting<br>for 17 year-olds")
```

```{r}
# Generate map for 2018
leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(-98.483330, 38.712046, zoom = 4) %>% 
  addPolygons(data = youth_vote_18, 
              fillColor = ~pal_register(youth_vote_18$primary_17), 
              fillOpacity = 0.7, 
              weight = 0.2, 
              smoothFactor = 0.2,
              popup = ~popup_register_18) %>%
  addLegend(pal = pal_register,
            values = youth_vote_18$primary_17,
            position = "bottomright", 
            title = "States with Primary Voting<br>for 17 year-olds")
```

